---
title: "Submissions Outcomes by Author"
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_4"
figure_extension <- ".pdf"

path_to_data <- "https://www.dropbox.com/s/u5lmsyri359mvr9/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(grid)
library(gridExtra)

# setup theme
theme_dakota <- function () { 
    theme_bw(base_size=10, base_family="Times") %+replace% 
        theme(
            strip.background = element_blank(),
            strip.text = element_text(size = 11),
            plot.title = element_text(size = 14, face = "bold"),
            axis.text.x = element_text(size = 12, margin = margin(t = 5, b = 5)),
            legend.title = element_blank(),
            legend.position = "bottom"
        )
}
s
# converts a provided p-value into a representation that is easier to visualize
sig2ast <- function(p) { ifelse(p <= 0.0001, "***", 
                                ifelse(p <= 0.001, "***", 
                                       ifelse(p <= 0.01, "**", 
                                              ifelse(p <= 0.05, "*", 
                                                     ifelse(p < 0.1, "•", "ns")
                                              )
                                       )
                                )
                          )
}

# I copied the below function from a stack overflow question. It should capitalize the first letter of each work in a string
# https://stackoverflow.com/questions/6364783/capitalize-the-first-letter-of-both-words-in-a-two-word-string
simpleCap <- function(x) {
  s <- strsplit(x, " ")[[1]]
  paste(toupper(substring(s, 1,1)), substring(s, 2),
      sep="", collapse=" ")
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
elife <- read_csv(path_to_data)
head(elife)
```

## Figure 4

> Figure 4. Percentage of full submissions that were accepted, shown by the gender of the corresponding author, first author, and last author. Authors whose gender is unknown are excluded from analysis. See Figure SI.2 for an extension of this figure including submission rates, encourage rates, and overall acceptance races. Vertical error bars indicate 95th confidence intervals of the proportion of submitted, encouraged, and accepted initial and full submissions. Asterisks indicate significance level of X2 tests of independence of frequency of encourage and acceptance by gender; “\*” = p < 0.05; “ns” = p > 0.05.

Note: There is a lot of extra code with this figure, because this figure originally showed informastion for submission rates and encourage and acceptance rates of initial submissions. Now we only show acceptance rates of full submissions. See figure SI.### for the expanded figure. 

```{r message=FALSE, warning=FALSE}
# First, I'll go through and run all the tests to get signif. values. I toyed around with ways to do this more programatically, but I think that this
# manual method is perhaps the simplest

# First, test the encourage rate diffs
s <- subset(elife, initial_decision_made)
t <- table("ca_gender" = s$ca_gender, "encouraged" = s$encouraged)[1:2,]
ca_encouraged <- prop.test(t[, 1:2])

# Difference in acceptance rates by gender of corresponding author
s <- subset(elife, initial_decision_made & (!has_full_submission | full_decision_made))
t <- table("ca_gender" = s$ca_gender, "accepted" = s$accepted)[1:2,]
ca_accepted <- prop.test(t[, 1:2])

# Difference in encourage-acceptance rates by gender of corresponding author
s <- subset(elife, initial_decision_made & full_decision_made)
t <- table("ca_gender" = s$ca_gender, "accepted" = s$accepted)[1:2,]
ca_ea <- prop.test(t[, 1:2])

# Difference in encourage-acceptance rates by gender of first author
s <- subset(elife, initial_decision_made & full_decision_made)
t <- table("fa_gender" = s$fa_gender, "accepted" = s$accepted)[1:2,]
fa_ea <- prop.test(t[, 1:2])

# Difference in encourage-acceptance rates by gender of last author
s <- subset(elife, initial_decision_made & full_decision_made)
t <- table("la_gender" = s$la_gender, "accepted" = s$accepted)[1:2,]
la_ea <- prop.test(t[, 1:2])

# Now we use the results of these tests to construct a table
measurements <- c(rep("submit_rate", 3), rep("encourage_rate", 3), rep("accept_rate", 3), rep("encourage_accept_rate", 3))
author <- rep(c("ca_gender", "fa_gender", "la_gender"), 4)

sig <- c("N/A", "N/A", "N/A", # submit rates, all NA since we don't really compare this
         sig2ast(ca_encouraged$p.value), "N/A", "N/A", # encourage rate
         sig2ast(ca_accepted$p.value), "N/A", "N/A", 
         sig2ast(ca_ea$p.value), sig2ast(fa_ea$p.value), sig2ast(la_ea$p.value)
         )
sig_df <- data.frame(measure_key = factor(measurements), author_key = factor(author), sig = sig)

# Now lets save these test results for later
sig_list <- list("Corresponding author enc. rate" = ca_encouraged, 
                   "Corresponding author acc. rate" = ca_accepted, 
                   "Corresponding author enc-acc rate" = ca_ea, 
                   "First author enc-acc rate" = fa_ea, 
                   "last author enc-acc rate" = la_ea)


data <- elife %>%
  filter(initial_decision_made) %>%
  gather(author_key, value = gender, ca_gender, la_gender, fa_gender) %>%
  filter(gender %in% c('M', 'F')) %>%
  group_by(author_key) %>%
  mutate(
    # if not ocnsidering the corr. author, then the count is all full submissions, ie: for when a submission ia made
    total_papers = ifelse(author_key == "ca_gender", n(), sum(has_full_submission))
    )%>%
  group_by(author_key, gender) %>%
  summarize(
        paper_count = n(),
        accept_count = sum(accepted == TRUE, na.rm = T),
        encourage_count = sum(encouraged == TRUE),
        encourage_rate = ifelse(author_key[1] == "ca_gender", encourage_count / n(), NA),
        # this accept rate is the number of accepted papers / number of initial + full submissions
        accept_rate = ifelse(author_key[1] == "ca_gender", accept_count / sum(!has_full_submission | full_decision_made, na.rm = T), NA),
        submit_rate = n() / total_papers[1],
        # this is the number of accepted papers divided by all papers 
        encourage_accept_rate = accept_count / sum(has_full_submission & full_decision_made, na.rm = T)
  ) %>%
  ungroup() %>%
  gather(measure_key, value = measurement, encourage_accept_rate) %>%
  left_join(sig_df, by = c("measure_key", "author_key")) %>%
  mutate(
    author_key = factor(author_key,
                        levels = c("ca_gender", "fa_gender", "la_gender"),
                        labels = c("ca_gender" = "Corresponding", "fa_gender" = "First", "la_gender" = "Last")
                        ),
    measure_key = factor(measure_key, 
                         labels = c("submit_rate" = "Submissions", "encourage_rate" = "Enc %", "accept_rate" = "Overall Acc %", "encourage_accept_rate" = "Accept %"),
                         levels = c("submit_rate", "encourage_rate", "accept_rate", "encourage_accept_rate")),
    gender = ifelse(gender == "F", "Female", "Male"),
    standard_error = 1.96 * sqrt( (measurement * (1 - measurement)) / paper_count),
    standard_error = standard_error * 100,
    measurement = measurement * 100,
    lower = measurement - standard_error,
    upper = measurement + standard_error
  )
  
  
gender_outcomes <- data %>%
  ggplot(aes(x = gender, y = measurement, color = factor(gender), shape = factor(gender), fill = factor(gender))) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, color = "black") +
  geom_text(aes(x = 1.5, y = 89, label = ifelse(measure_key == "Submissions", "", as.character(sig))), color = "black", size = 4) +
  geom_segment(aes(x = 1, xend = 2, y = ifelse(sig != "N/A", 85, 1.1), yend = ifelse(sig != "N/A", 85, 1.1)), color = "black") +
  facet_wrap(~author_key) +
  coord_cartesian(ylim = c(0, 100)) +
  theme_dakota() +
  guides(fill = F) +
  labs(
    x = "Gender of Author",
    y = "% of full submissions accepted",
    title = "A"
  ) +
  theme(plot.title = element_text(size = 20),
        strip.text = element_text(size = 11),
        axis.title.x = element_blank(),
        axis.text.x = element_text(size = 10, face = "bold"),
        axis.title.y = element_text(size = 12))

gender_outcomes
```




## Misc. information

Signif. test results for figure 4
```{r}
fig4_sig_table <- data.frame(
  test = names(sig_list),
  p.value = sapply(sig_list, function(x) x$p.value),
  estimate = sapply(sig_list, function(x) abs(x$estimate[1] - x$estimate[2])),
  n = sapply(sig_list, function(x) sum(x$observed)),
  statistic = sapply(sig_list, function(x) x$statistic),
  #df = sapply(sig_list, function(x) x$parameter)
  lower = sapply(sig_list, function(x) x$conf.int[1]),
  upper = sapply(sig_list, function(x) x$conf.int[2])
)

rownames(fig4_sig_table)<- NULL
fig4_sig_table
```

### Data used to produce figure 4
```{r rows.print=15}
data %>%
  select(-sig, -lower, -upper, -standard_error) %>%
  spread(measure_key, measurement)
```

## Gender distirbution of gender-identified authors on full submissions at eLife, at manuscript level, includes duplicate authors
```{r}
full <- elife %>% 
  filter(has_full_submission)
```

Of correpsonding authors
```{r}
full %>% 
  filter(ca_gender %in% c("M", "F")) %>%
  group_by(ca_gender) %>% 
  summarize(
    freq = n(), 
    prop = freq / dim(full)[1]
  )
```

First authors...
```{r}
full %>% 
  filter(fa_gender %in% c("M", "F")) %>%
  group_by(fa_gender) %>% 
  summarize(
    freq = n(), 
    prop = freq / dim(full)[1]
  )
```

And last authors
```{r}
full %>% 
  filter(la_gender %in% c("M", "F")) %>%
  group_by(la_gender) %>% 
  summarize(
    freq = n(), 
    prop = freq / dim(full)[1]
  )
```


## Submissions and outcomes by country

> Figure 6. Top: proportion of all initial submissions, encouraged initial submissions, and accepted full submissions comprised by the national affiliation of the corresponding author for the top eight most prolific countries in terms of initial submissions. Bottom: Encourage rate of initial submissions, acceptance rate of initial submissions, and acceptance rate of full submissions by national affiliation of the corresponding author for the top eight more prolific countries in terms of initial submissions. Error bars on bottom panel indicate standard error of proportion of encouraged initial submissions and accepted initial and full submissions for each country. This same graph with the top 16 most prolific nations can be found in Figure SI.4

```{r message=FALSE, warning=FALSE}
# Its easiest and most efficient to calculate these total values outside of the dplyr chain below
total_submissions <- length(unique(elife$MSNO))
total_full_submissions <- sum(!elife$full_decision_date_isNA, na.rm = T)
total_encouraged <- with(elife, sum(initial_decision_made & encouraged, na.rm = T))
total_accepted <- with(elife, sum(full_decision_made & accepted == T, na.rm = T))

# Firs the first plot, this one will show the proportion that each countrie constitutes the total number of submitted, encouraged, and accepted manuscripts
data_national_p1 <- elife %>%
  filter(initial_decision_made & (full_decision_made | full_decision_date_isNA) & !is.na(ca_country)) %>%
  mutate(ca_country = factor(sapply(as.character(ca_country), simpleCap))) %>% # capitalize first letter of country stirngs
  group_by(ca_country) %>%
  summarize(
    paper_count = n(),
    accept_count = sum(accepted == TRUE, na.rm = T),
    encourage_count = sum(encouraged == TRUE, na.rm = T),
    full_submissions = sum(has_full_submission),
    prop_of_full =  full_submissions / total_full_submissions,
    prop_of_initial = paper_count / total_submissions,
    prop_of_encouraged = encourage_count / total_encouraged,
    prop_of_accepted = accept_count / total_accepted
  ) %>%
  group_by(ca_country) %>%
  mutate(totals = sum(paper_count)) %>%
  ungroup() %>%
  top_n(8, totals) %>%
  gather(prop_key, value = prop, prop_of_initial, prop_of_full, prop_of_accepted) %>%
  mutate(prop_key = factor(prop_key, 
                           levels = c("prop_of_initial", "prop_of_full", "prop_of_accepted"),
                           labels = c("prop_of_initial" = "Prop. of initial sub.", 
                                      "prop_of_full" = "Prop. of full sub.", 
                                      "prop_of_accepted" = "Prop. of accepted sub.")),
         ca_country = reorder(ca_country, paper_count),
         prop = prop * 100
  ) 

p1 <- data_national_p1 %>%
  # Start ggplot
  ggplot(aes(x = ca_country, y = prop)) +
  geom_bar(stat = "identity", color = "black", alpha = 0.8, fill = "dodgerblue4", position = position_dodge(1)) +
  geom_text(aes(y = prop, label = paste0(round(prop, 1), "%")), size = 3, hjust = -0.2, family = "Times", position = position_dodge(1)) +
  scale_y_continuous(limits = c(0, 1)) +
  facet_wrap(~prop_key, labeller = labeller(key = labels)) +
  theme_dakota() +
  coord_flip() +
  ylim(0, 75) +
  labs(y = "",
       title = "B") +
  theme(
    strip.background = element_blank(),
    axis.title.x = element_blank(),
    text = element_text(size = 12, family = "Times"),
    plot.title = element_text(size = 20),
    axis.text.x = element_text(size = 9),
    axis.title.y = element_blank()
  )
  

data_national_p2 <- elife %>%
  filter(initial_decision_made & (full_decision_made | full_decision_date_isNA) & !is.na(ca_country)) %>%
  mutate(ca_country = factor(sapply(as.character(ca_country), simpleCap))) %>% # capitalize first letter of country stirngs
  group_by(ca_country) %>%
  summarize(
    paper_count = n(),
    accept_count = sum(accepted == TRUE, na.rm = T),
    encourage_count = sum(encouraged == TRUE, na.rm = T),
    encourage_rate = encourage_count / paper_count,
    accept_rate = accept_count / sum(!has_full_submission | full_decision_made, na.rm = T),
    encourage_accept_rate = accept_count / sum(has_full_submission & full_decision_made, na.rm = T)
  ) %>%
  group_by(ca_country) %>%
  mutate(totals = sum(paper_count)) %>%
  ungroup() %>%
  top_n(8, totals) %>%
  mutate(
    ca_country = reorder(ca_country, encourage_rate)
  ) %>%
  gather(measure_key, value = measurement, encourage_rate, accept_rate, encourage_accept_rate) %>%
  mutate(measure_key = factor(measure_key, 
                           levels = c("encourage_rate", "encourage_accept_rate", "accept_rate"),
                           labels = c("encourage_rate" = "% of initial sub. encouraged", 
                                      "encourage_accept_rate" = "% of full sub. accepted",
                                      "accept_rate" = "% of initial sub. accepted"
                                      ))
  ) %>%
  group_by(measure_key) %>%
  mutate(
    se = 1.96 * sqrt(measurement * (1 - measurement) / paper_count),
    measurement = measurement * 100,
    se = se * 100,
    lower = measurement - se,
    upper = measurement + se
  ) 

p2 <- data_national_p2 %>%
  ggplot(aes(x = ca_country, y = measurement)) +
  geom_bar(stat = "identity", color = "black", fill = "dodgerblue4", alpha = 0.8) +
  geom_text(aes(y = measurement, label = paste0(round(measurement, 1), "%")), size = 3, hjust = -0.6, family = "Times") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  facet_wrap(~measure_key, labeller = labeller(key = labels)) +
  theme_dakota() +
  labs(y = "%") +
  coord_flip() +
  ylim(0, 75) +
  theme(
    axis.title.y = element_blank(),
    axis.title.x element_blank(),
    strip.background = element_blank(),
    text = element_text(size = 12, family = "Times"),
    axis.text.x = element_text(size = 9),
    plot.title = element_text(size = 20)
  ) +
  labs(title = "C")

national_outcomes <- grid.arrange(p1, p2)

national_outcomes
```


### Data to produce figure 5, the top panel
```{r}
data_table <- data_national_p1 %>%
  spread(prop_key, prop)

data_table
```

### Proportion of intiial submissions constituted by top 8 counties
```{r}
sum(data_table$`Prop. of initial submissions`)
```

### Proportion of full submissions constituted by top 8 counties
```{r}
sum(data_table$`Prop. of full submissions`)
```

### Proportion of accepted submissions constituted by top 8 counties
```{r}
sum(data_table$`Prop. of accepted submissions`)
```

###Data used to produce figure 5, bottom panel
```{r rows.print = 24}
data_table <- data_national_p2 %>%
  select(-se, -lower, -upper)

data_table
```

## Combine figures

Construct a combined figure
```{r}
g <- grid.arrange(gender_outcomes, national_outcomes, ncol = 2, widths = 1:2)
plot(g)
```


## Save the file
```{r}
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/main/", figure_name, figure_extension)
ggsave(name, plot = g, height = 5, width = 10.5)
```


