---
title: "Gatekeeper Representation"
author: "Dakota Murray"
output: html_notebook
---

### Set parameters for the name and extension of the figure
```{r}
figure_name <- "fig_3"
figure_extension <- ".pdf"

path_to_data <- "https://www.dropbox.com/s/u5lmsyri359mvr9/formatted_elife.csv?dl=1"
```

```{r message=FALSE, warning=FALSE}
# Load required packages
library(tidyverse)
library(grid)
library(gridExtra)

# setup theme
theme_dakota <- function () { 
    theme_bw(base_size=10, base_family="Times") %+replace% 
        theme(
            strip.background = element_blank(),
            strip.text = element_text(face = "bold"),
            plot.title = element_text(size = 11, face = "bold"),
            axis.text.x = element_text(size = 9, margin = margin(t = 5, b = 5)),
            legend.title = element_blank(),
            legend.position = "bottom"
        )
}

# converts a provided p-value into a representation that is easier to visualize
sig2ast <- function(p) { ifelse(p <= 0.0001, "***", 
                                ifelse(p <= 0.001, "***", 
                                       ifelse(p <= 0.01, "**", 
                                              ifelse(p <= 0.05, "*", 
                                                     ifelse(p < 0.1, ".", "ns")
                                              )
                                       )
                                )
                          )
}
```

### Load the required data
```{r message=FALSE, warning=FALSE}
people <- read_csv("https://www.dropbox.com/s/ic2b17xp3wo0255/elife_people.csv?dl=1")
head(people)
```


### Plot the figure

>Figure 3. Top: proportion of identified men and women in the populations of distinct gatekeepers (senior editors, reviewing editors, and peer reviewers) and of the populations of distinct corresponding authors, first authors, and last authors; percentages exclude those for whom gender could be identified. Bottom: proportion of people with national affiliations within each of six continents in the population of distinct gatekeepers, and for the population of distinct corresponding, first, and last authors. Black dashed lines overlaid on authorship graphs indicate the proportion of gatekeepers within that gendered or continental category. Asterisks indicate the significance level of X2 tests of independence comparing the frequency of gender or continents between gatekeepers and each authorship type. “\*\*\*\*“ = p < 0.0001; “ns” = p > 0.05.

```{r message=FALSE, warning=FALSE}
# Now lets get the significance levels...
s <- subset(people, type == "Corr. Author" | type == "Gatekeeper" )
t <- t(table(s$gender, s$type))
corr <- chisq.test(t[1, 1:2], p = prop.table(t[2, 1:2]))


s <- subset(people, type == "First Author" | type == "Gatekeeper" )
t <- t(table(s$gender, s$type))
first <- chisq.test(t[1, 1:2], p = prop.table(t[2, 1:2]))


s <- subset(people, type == "Last Author" | type == "Gatekeeper" )
t <- t(table(s$gender, s$type))
last <- chisq.test(t[2, 1:2], p = prop.table(t[1, 1:2]))

sig_df <- data.frame(type = rep(c("Gatekeeper", "Corr. Author", "First Author", "Last Author"), 1), 
                     gender = rep(c("Male", "Female"), 2),  
                     sig = c("N/A", sig2ast(corr$p.value), sig2ast(first$p.value), sig2ast(last$p.value)))

gender_sig_list <- list("Corresponding Author" = corr, "First Author" = first, "Last Author" = last)


expected <- people %>%
  filter(gender %in% c("M", "F")) %>%
  mutate(gender = ifelse(gender == "M", "Male", "Female")) %>%
  filter(type == "Gatekeeper") %>%
  mutate(total = n()) %>%
  group_by(type, gender) %>%
  summarize(
    expected_prop = n() / total[1]
  ) %>%
  ungroup() %>%
  select(-type)


p1 <- people %>%
  filter(gender %in% c("M", "F")) %>%
  mutate(gender = ifelse(gender == "M", "Male", "Female")) %>%
  group_by(type) %>%
  mutate(total = n()) %>%
  group_by(type, gender) %>%
  summarize(
    prop = n() / total[1]
  ) %>%
  ungroup() %>%
  left_join(expected, by = "gender") %>%
  left_join(sig_df, by = c("gender", "type")) %>%
  mutate(
    type = factor(type, 
                  levels = c("Gatekeeper", "Corr. Author", "First Author", "Last Author"),
                  labels = c("Gatekeeper" = "Gatekeeper", "Corr. Author" = "Corresponding", "First Author" = "First", "Last Author" = "Last")),
    sig = as.character(sig)
  ) %>%
  # start ggplot
  ggplot(aes(x = gender, y = prop, fill = gender)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  geom_bar(stat = "identity", aes(y = expected_prop), alpha = 0, color = "black", fill = "white", size = 0.5, linetype = "longdash") +
  #geom_text(aes(label = paste0(round(prop, 3) * 100, "%"), y = 0.1), size = 3) +
  geom_text(aes(label = ifelse(sig == "N/A", "", sig), x = 1.5, y = 0.9)) +
  geom_segment(aes(x = 1, xend = 2, y = ifelse(sig != "N/A", 0.85, 1.1), yend = ifelse(sig != "N/A", 0.85, 1.1)), color = "black") +
  facet_wrap(~type, nrow = 1) +
  theme_dakota() +
  ylim(c(0, 1)) +
  theme(
    axis.title.x = element_blank(),
    legend.position = "right",
    plot.title = element_text(size = 18)
  ) +
  labs(y = "Proportion",
       title = "A") +
  guides(fill = F)

s <- subset(people, continent != "Antarctica" & (type == "Corr. Author" | type == "Gatekeeper" ))
t <- t(table(s$continent, s$type))
corr <- chisq.test(t[1, ], p = prop.table(t[2, ]))

s <- subset(people, continent != "Antarctica" & (type == "First Author" | type == "Gatekeeper" ))
t <- t(table(s$continent, s$type))
first <- chisq.test(t[1, ], p = prop.table(t[2, ]))

s <- subset(people, continent != "Antarctica" & (type == "Last Author" | type == "Gatekeeper" ))
t <- t(table(s$continent, s$type))
last <- chisq.test(t[2, ], p = prop.table(t[1, ]))

sig_df <- data.frame(type = rep(c("Gatekeeper", "Corr. Author", "First Author", "Last Author"), 3), 
                     continent = rep(c("Africa", "Asia", "Europe", "North America", "Oceania", "South America"), 2),  
                     sig = c("N/A", sig2ast(corr$p.value), sig2ast(first$p.value), sig2ast(last$p.value)))

continent_sig_list <- list("Corresponding Author" = corr, "First Author" = first, "Last Author" = last)

# Now we should repeat this graph, except this time for the continental representation
expected <- people %>%
  filter(type == "Gatekeeper") %>%
  mutate(total = n()) %>%
  filter(continent != "Antarctica") %>%
  group_by(type, continent) %>%
  summarize(
    expected_prop = n() / total[1]
  ) %>%
  ungroup() %>%
  select(-type)


p2 <- people %>%
  group_by(type) %>%
  mutate(total = n()) %>%
  filter(continent != "Antarctica") %>%
  group_by(type, continent) %>%
  summarize(
    prop = n() / total[1]
  ) %>%
  ungroup() %>%
  left_join(expected, by = "continent") %>%
  left_join(sig_df, by = c("continent", "type")) %>%
  mutate(
    type = factor(type, 
                  levels = c("Gatekeeper", "Corr. Author", "First Author", "Last Author"),
                  labels = c("Gatekeeper" = "Gatekeeper", "Corr. Author" = "Corresponding", "First Author" = "First", "Last Author" = "Last")),
    continent = reorder(continent, -prop),
    sig = as.character(sig)
  ) %>%
  # start ggplot
  ggplot(aes(x = continent, y = prop, fill = continent)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  geom_bar(stat = "identity", aes(y = expected_prop), alpha = 0, color = "black", fill = "white", size = 0.5, linetype = "longdash") +
  #geom_text(aes(label = paste0(round(prop, 3) * 100, "%"), y = 0.05), size = 2) +
  geom_text(aes(label = ifelse(sig == "N/A", "", sig), x = 3.5, y = 0.92)) +
  geom_segment(aes(x = 1, xend = 6, y = ifelse(sig != "N/A", 0.85, 1.1), yend = ifelse(sig != "N/A", 0.85, 1.1)), color = "black") +
  facet_wrap(~type, nrow = 1) +
  theme_dakota() +
  ylim(c(0, 1)) +
  guides(fill = F) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.x = element_blank(),
    plot.title = element_text(size = 18)
  ) +
  labs(y = "Proportion",
       title = "B") +
  scale_fill_brewer(palette = "Set2")

gate_rep <- grid.arrange(p1, p2)
gate_rep
```


## Save the files
```{r}
name <- paste0("~/Dropbox/eLife/elife-analysis/figures/main/", figure_name, figure_extension)
ggsave(name, plot = gate_rep, height = 6, width = 7)
```


And below are the explicit results of the significaince tests used in the previous graphs to compare the actual proportion of male/female authors with the expected proportion given the population of gatekeepers. 
```{r}
fig3_top_sig_table <- data.frame(
  authorship = c("Corresponding", "First", "Last"),
  p.value = sapply(gender_sig_list, function(x) x$p.value),
  n = sapply(gender_sig_list, function(x) sum(x$observed)),
  statistic = sapply(gender_sig_list, function(x) x$statistic),
  df = sapply(gender_sig_list, function(x) x$parameter)
)

rownames(fig3_top_sig_table) <- NULL
fig3_top_sig_table
```

And the same for the continent-level tests
```{r}
fig3_bottom_sig_table <- data.frame(
  continent = c("Corresponding", "First", "Last"),
  p.value = sapply(continent_sig_list, function(x) x$p.value),
  n = sapply(continent_sig_list, function(x) sum(x$observed)),
  statistic = sapply(continent_sig_list, function(x) x$statistic),
  df = sapply(continent_sig_list, function(x) x$parameter)
)

rownames(fig3_bottom_sig_table) <- NULL
fig3_bottom_sig_table
```

### Frequencies of author and gatekeeper genders at eLife
```{r rows.print=15}
library(stargazer)
gender_dem <- people %>% 
  group_by(type) %>%
  mutate(total = n()) %>%
  group_by(type, gender) %>%
  summarize(
    freq = n(),
    prop = round(freq / total[1], 3)
  )

names(gender_dem) <- c("Role", "Gender", "#", "%")

stargazer(gender_dem, summary = F, rownames = F, type = "latex", font.size = "scriptsize")


```

### Frequencies of author and gatekepeer continental affilaitions at eLife
```{r rows.print=30}
geo_dem <- people %>% 
  group_by(type) %>%
  mutate(total = n()) %>%
  group_by(type, continent) %>%
  summarize(
    freq = n(),
    prop = round(freq / total[1], 3)
    ) %>%
  mutate(continent = ifelse(is.na(continent), "UNK", continent))

names(geo_dem) <- c("Role", "Continent", "#", "%")

stargazer(geo_dem, summary = F, rownames = F, type = "latex", font.size = "scriptsize")
```

